On multilevel constraint decomposition methods \\ for nonlinear variational inequalities

### 2021-07-13 paper/mcd2.tex
Use float package to manage pseudocode formatting better.
A cartoon of constraint decomposition.

\usepackage[within=section]{newfloat}

@@

@@

\DeclareFloatingEnvironment[name=Pseudocode]{pcode}

@@

@@

Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.  For a continuous \emph{residual functional} $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$:
\begin{equation}
\ip{f(u^*)}{v-u^*} \ge \ip{g}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:vi}
\end{equation}
Because $f$ is a (generally) nonlinear map, $g$ is not strictly needed, but its presence is helpful to the algorithms of Section \ref{sec:multilevel}.  (By redefining $f$ we may take $g=0$.)

@@

@@

We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$.  (By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.  Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.  If \pr{parallel} $=$ \text{\textbf{true}} then the \text{\textbf{for}} loop can be computed in any order.

\begin{pseudo*}
\pr{cd}(0<\alpha\le 1, u\in\cK)\text{:} \\+
    if \pr{parallel}: \\+
        for $i \in \{0,\dots,m-1\}$: \\+
            $\hat w_i\in \cK_i$: \\+
                 $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\---
    else: \\+
        for $i = 0,\dots,m-1$: \\+
            $\hat w_i\in \cK_i$: \\+
                $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
            $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\--
    $\hat w = \sum_i \hat w_i$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}

>>>>

@@

\usepackage{float}

@@

@@

\newfloat{pseudofloat}{t}{xyz}[section]
\floatname{pseudofloat}{Algorithm}

@@

Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.  For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$:
\begin{equation}
\ip{f(u^*)}{v-u^*} \ge \ip{g}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:vi}
\end{equation}
Because $f$ is a (generally) nonlinear map, $g$ is not strictly needed, but its presence is helpful to the algorithms of Section \ref{sec:multilevel}.  (By redefining $f$ we may take $g=0$.)

@@

\begin{figure}
\includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
\caption{A constraint decomposition (CD) in a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ looks something like this.}
\label{fig:cartoon}
\end{figure}

@@

We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$ (Algorithm \ref{alg:basiccd}).  (By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.  Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.  If \pr{parallel} $=$ \text{\textbf{true}} then the \text{\textbf{for}} loop can be computed in any order.

\begin{pseudofloat}
\begin{pseudo*}
\pr{cd}(0<\alpha\le 1, u\in\cK)\text{:} \\+
    if \pr{parallel}: \\+
        for $i \in \{0,\dots,m-1\}$: \\+
            $\hat w_i\in \cK_i$: \\+
                 $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\---
    else: \\+
        for $i = 0,\dots,m-1$: \\+
            $\hat w_i\in \cK_i$: \\+
                $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
            $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\--
    $\hat w = \sum_i \hat w_i$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}
\caption{The basic constraint decomposition (CD) algorithm for VI problem \eqref{eq:vi}.}
\label{alg:basiccd}
\end{pseudofloat}
